Wolbachia, Toxoplasma and O. Cordyceps: Extreme Host-Manipulating Microbes That Changed the World

### 2015-12-16 paper.tex
Added some impacts / analysis

\textit{Ophiocordyceps unilateralis} was first discovered in 1859 by the British naturalist Alfred Russel Wallace who, probably not coincidently, is best known for independently conceiving the theory of evolution through natural selection. Darwin published his now famous "On the Origin of Species" in the same year, which gives some good context as to where the scientific paradigms were at the time.\cite{darwin}

@@

- Cordyceps has been studied to make an immuno-surpressant drug
A cousin of the and manipulating fungus \textit{Cordyceps sinensis} \cite{medicinal_cordy}
- Cordyceps is necessary for keeping bug populations in check
- Toxoplasma has been correlated to a rise in schizophernia
- Wolbachia has made us question evolution
- Infecting insects with Wolbachia causes them to not carry dengue fever
- Targetting Wolbachia to treat filarial infections is effective.

>>>>

\textit{Ophiocordyceps unilateralis} was first discovered by the British naturalist Alfred Russel Wallace who, probably not coincidently, is best known for independently conceiving the theory of evolution through natural selection.

@@

- Cordyceps has been studied to make an immuno-surpressant drug

- Cordyceps is necessary for keeping bug populations in check

- Toxoplasma has been correlated to a rise in schizophernia

- Wolbachia has made us question evolution

The presence of \textit{Wolbachia} in filarial nematodes has raised interesting questions about how \textit{Wolbachia} evolves in its host species. Some scientists discovered that \textit{Wolbachia} "have possibly been lost during evolution along some lineages of filarial nematodes."\cite{Wevolution_loss} The reason this discovery is bizare is because \textit{Wolbachia}'s ability to evolve into the life-cycles of so many different species makes sense if \textit{Wolbachia} gives the infected a fitness advantage over uninfected members. However, if species evolve to get rid of \textit{Wolbachia} then this points to the fact that \textit{Wolbachia} does not supply them with a fitness advantage.

So why would one species derive an advantage from a \textit{Wolbachia} infection whereas another species might not? The answer is that nobody really knows, at least not yet. The impact this discovery has had on our understanding of evolution has not truly been felt yet. However, when the reason for differences in derived fitness advantage from \textit{Wolbachia} infections is understood it could uncover a greater understanding of how interactions between organic matter works. It might explain why some people like coffee and find it helps them get work done, while others do not even like the smell of coffee.

- Infecting insects with Wolbachia causes them to not carry dengue fever

Aside from helping us understand the world better, \textit{Wolbachia} may have the ability to help cure humans of deadly diseases. A recent study found that \textit{Aedes aegypti} mosquitos infected with \textit{Wolbachia} cause it to no longer be able to carry the pathogen that causes Dengue Fever.\cite{Wdengue_fever}

This is big news, because it might lead to the eradication of Dengue Fever, for which there is currently no vaccine. By artificially bringing \textit{Wolbachia} into the \textit{Aedes aegypti} species the fear of Dengue Fever could vanish entirely. Not only does this impact those in Dengue Fever risk areas, but it may also lead to the prevention of other diseases as well. Many scientists are looking into the idea of infecting mosquitos that harbor \textit{Milaria} with \textit{Wolbachia} as a prevention technique.\cite{Wmilaria}

Using \textit{Wolbachia} as an "antidote" for different diseases is novel and has potential to change the world dramatically. The best part is that it is a one-time use antidote, because \textit{Wolbachia} has the tools to infect an entire species through vertical transmission. It may also cause scientists to look at other symbiotic relationships between organisms, and see if they have any potential to help humans. There might be a new era of using living organisms as tools to cure diseases and help humans in other ways.

- Targetting Wolbachia to treat filarial infections is effective.

The usage of the relationship between \textit{Wolbachia} and its host may even be abused in other ways.
